---
title: "manipulate"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(DT)
```

First we're going to read in the California current data from the ERDDAP server
https://oceanview.pfeg.noaa.gov/erddap/index.html
could use ERDDAP to filter and select but we're going to use the most basic version so that we can filter it ourselves.

## Read online table, download to (`*.csv`)

```{r}
# set variables
csv_url  <- "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_AC.csv"
dir_data <- "data"

# derived variables
csv <- file.path(dir_data, basename(csv_url))

# create directory
#only need to do once
#dir.create(dir_data)

# download file
if (!file.exists(csv))
  download.file(csv_url, csv)
```
### Read table
```{r}
# attempt to read csv
d <- read.csv(csv)

# show the data frame
d


# read csv by skipping first two lines, so no header
d <- read.csv(csv, skip = 2, header = FALSE)
d

#tibble::tibble(d)
#glimpse(d)

# update data frame to original column names from the file
#this is necessary because you skipped two lines which included header
names(d) <- names(read.csv(csv))
d
```
## show table `DT::datatable()`

This makes the table interactive in the knit html, and displays better than any of the previous view options.


```{r}
#show table
DT::datatable(d)

#or

library(DT)
datatable(d)
```

## Wrangle data

### Manipulate with `dplyr`

To do:
convert df to tibble
transform time column to as.Date()
select specific columns
filter by row values

We only want to see information relating to revenues in a specific year

```{r}
library(DT)
library(dplyr)

d <- d %>% 
  # tibble
  tibble() %>% 
  # mutate time
  mutate(
    time = as.Date(substr(time, 1, 10))) %>% 
  # select columns
  select(
    time, 
    starts_with("total_fisheries_revenue")) %>% 
  # filter rows
  filter(
    time >= as.Date("1981-01-01"))

datatable(d)
```

### Tidy with `tidyr`

Here we're getting long data for summaries in future sections.

```{r}
library(tidyr)

d <- d %>% 
  pivot_longer(-time)

datatable(d)
```

### Summarize with `dplyr`

our pipe will also use stringr to adjust row values to only contain state codes

```{r}
library(stringr)

#state codes only
d <- d %>% 
  mutate(
    region = str_replace(name, "total_fisheries_revenue_", "")) %>% 
  select(time, region, value)
datatable(d)


#make a datatable to summarize the average revenue across the coast and by state
d_sum <- d %>% 
  group_by(region) %>% 
  summarize(
    avg_revenue = mean(value))
datatable(d_sum) %>% 
  formatCurrency("avg_revenue")

```

## Apply functions with `purrr` on a `nest`ed `tibble`

Here we're taking advantage of the fact that tibbles can nest to create a tibble that has data grouped by region

```{r}
#group by region
n <- d %>% 
  group_by(region) %>% 
  nest(
    data = c(time, value))
n

#run a linear model on each tibble nested in the tibble
n <- n %>% 
  mutate(
    lm    = map(data, function(d){
      lm(value ~ time, d) } ),
    trend = map_dbl(lm, function(m){
      coef(summary(m))["time","Estimate"] }))
n

#now redefine the tibble to jsut be the results of the linear model
n %>% 
  select(region, trend) %>% 
  datatable()
```

## rmarkdown notes

If you’d like to render your Rmarkdown into HTML with a table of contents (toc), showing up to 3rd level headings (i.e. ### Heading; use toc_depth: 3) and have a floating menu (toc_float), along with the ability to Show (or Hide) your Code Chunks (code_folding), then replace this default front matter yaml:

output: html_document
with:

output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: "show"
For details, see 3.1 HTML document | R Markdown: The Definitive Guide. : https://bookdown.org/yihui/rmarkdown/html-document.html



## Further learning objectives...

try some stuff on the list here; https://noaa-iea.github.io/r3-train/manipulate.html#Further_Learning_Objectives

