---
title: "visualize"
output: html_document
---

Learning objectives; use of here(), ggplot2 (grammar of graphics), interactive online plots using htmlwidgets

## Read in data

this time we will read directly from csv either from erdap server or the csv we downloaded

note, read_csv vs read.csv ; read_csv() does not automatically convert characters to factors and defaults to tibbles instead of dataframes.

```{r}
# libraries
library(here)
library(readr)
library(DT)

# variables
url_ac <- "https://oceanview.pfeg.noaa.gov/erddap/tabledap/cciea_AC.csv"
# if ERDDAP server down (Error in download.file) with URL above, use this:
#    url_ac <- "https://raw.githubusercontent.com/noaa-iea/r3-train/master/data/cciea_AC.csv"
csv_ac <- here("data/cciea_AC.csv")

# download data
if (!file.exists(csv_ac))
  download.file(url_ac, csv_ac)

# read data
d_ac        <- read_csv(csv_ac, col_names = F, skip = 2)
names(d_ac) <- names(read_csv(csv_ac))

# show data
datatable(d_ac)
```
## Visualize

### line plot

First must format with dplyr. then specify in aesthetics and add geometry.
```{r}
library(dplyr)
library(ggplot2)

# subset data
d_coast <- d_ac %>% 
  # select columns
  select(time, total_fisheries_revenue_coastwide) %>% 
  # filter rows
  filter(!is.na(total_fisheries_revenue_coastwide))
datatable(d_coast)

# ggplot object
p_coast <- d_coast %>% 
  # setup aesthetics
  ggplot(aes(x = time, y = total_fisheries_revenue_coastwide)) +
  # add geometry
  geom_line()

# show plot
p_coast

#add trend line

p_coast + 
  geom_smooth(method = "lm")

```

### histogram

adjust binwidth as needed

```{r}
d_coast %>% 
  # setup aesthetics
  ggplot(aes(x = total_fisheries_revenue_coastwide)) +
  # add geometry
  geom_histogram(binwidth = 10)
```

### series line plot 

color is region here. We need to tidy data into long format as in the manipulate lesson.


```{r}
library(stringr)
library(tidyr)

## this is the stuff from last time

d_rgn <- d_ac %>% 
  # select columns
  select(
    time, 
    starts_with("total_fisheries_revenue")) %>% 
  # exclude column
  select(-total_fisheries_revenue_coastwide) %>% 
  # pivot longer
  pivot_longer(-time) %>% 
  # mutate region by stripping other
  mutate(
    region = name %>% 
      str_replace("total_fisheries_revenue_", "") %>% 
      str_to_upper()) %>% 
  # filter for not NA
  filter(!is.na(value)) %>% 
  # select columns
  select(time, region, value)

## this is the plot
  
# create plot object
p_rgn <- ggplot(
  d_rgn,
  # aesthetics
  aes(
    x     = time, 
    y     = value,
    group = region,
    color = region)) +
  # geometry
  geom_line()

# show plot
p_rgn

#updating the labels
p_rgn <- p_rgn +
  labs(
    title  = "Fisheries Revenue",
    x      = "Year",
    y      = "Millions $ (year 2015)",
    color  = "Region")
p_rgn

```
### facet wrap for multiple plots

Instead of putting all the lines on top of each other here we have them in separate plots by region. not an optimal example for actual usage but works for illustrating the point.
```{r}
p_rgn +
  facet_wrap(vars(region))
```
### Bar plot

geom_bar() and geom_col() are different; bar only uses a discrete x but col uses a discrete x and continuous y

Here we are using geom_col to compare each region's revenue for a single year.

Note the use of glue to make the code reusable for multiple years or in future years. The object yr_max is pasted into a string instead of being written out each time.

```{r}
library(glue)
library(lubridate)

yr_max <- year(max(d_rgn$time))

d_rgn %>% 
  # filter by most recent time
  filter(year(time) == yr_max) %>% 
  # setup aesthetics
  ggplot(aes(x = region, y = value, fill = region)) +
  # add geometry
  geom_col() +
  # add labels
  labs(
    title  = glue("Fisheries Revenue for {yr_max}"),
    x      = "Region",
    y      = "Millions $ (year 2015)",
    fill   = "Region")
```
### variation of a series, boxplot

```{r}
d_rgn %>% 
  # setup aesthetics
  ggplot(aes(x = region, y = value, fill = region)) +
  # add geometry
  geom_boxplot() +
  # add labels
  labs(
    title  = "Fisheries Revenue Variability",
    x      = "Region",
    y      = "Millions $ (year 2015)") +
  # drop legend since redundant with x axis
  theme(
    legend.position = "none")
```
### Variation of a series, violin

```{r}
p_rgn_violin <- d_rgn %>% 
  # setup aesthetics
  ggplot(aes(x = region, y = value, fill = region)) +
  # add geometry
  geom_violin() +
  # add labels
  labs(
    title  = "Fisheries Revenue Variability",
    x      = "Region",
    y      = "Millions $ (year 2015)") +
  # drop legend since redundant with x axis
  theme(
    legend.position = "none")
p_rgn_violin
```
## Change theme

```{r}
p_rgn_violin +
  theme_classic()
```

## Interactive plots with plotly or dygraphs

### plotly

One advantage here is that you can use your cursor on the plot to see any given datapoint, turn on and off items in the legend, download a png of the adjusted plot, etc.

Not as reproducible for final figures of course but good for what it is.

```{r}
plotly::ggplotly(p_rgn)
```

### interactive time series with dygraphs

Here you use wide data to make the plot, and you can scroll along the time series.

```{r}
library(dygraphs)

#first make wide data

d_rgn_wide <- d_rgn %>% 
  mutate(
    Year = year(time)) %>% 
  select(Year, region, value) %>% 
  pivot_wider(
    names_from  = region,
    values_from = value)
datatable(d_rgn_wide)


d_rgn_wide %>% 
  dygraph() %>% 
  dyRangeSelector()
```


