---
title: "04_map"
output: html_document
---

Mapping!

note; the original code here using require() made my r really mad
```{r}
library(librarian)
# librarian::shelf() is like library() except installs package if missing, 
#   even from Github if include owner/repo
shelf(
  NOAA-EDAB/ecodata,
  sf)

```

## Get data

epu_sf are eastern US/CA ocean ecoregion data
```{r}
library(ecodata)
library(dplyr)

ecodata::epu_sf

epu_sf <- ecodata::epu_sf %>% 
  st_transform(4326)

class(epu_sf)

# "sf"         "data.frame"
g1 <- epu_sf$geometry[1]
# see in Environment pane, expand g1

plot(epu_sf)

plot(epu_sf["EPU"])
```
```{r}
shelf(mapview)

#this lets you scroll in and oout and see it on a map
mapview(epu_sf)
```


```{r}
shelf(leaflet)

leaflet() %>% 
  #addTiles() %>%
  addProviderTiles(providers$Esri.OceanBasemap) %>% 
  addPolygons(data = epu_sf)
```
### Group by, extract from erddap

erddap
https://coastwatch.pfeg.noaa.gov/erddap/griddap/jplMURSST41mday.html

```{r}
shelf(
  here,
  rerddap)

#character string to cache erddap result
sst_gd_rds <- here("data/sst_gd.rds")

#this is from sf package 
epu_bb <- st_bbox(epu_sf)
epu_bb

#name is from url to erddap data
sst_info <- info('jplMURSST41mday')
sst_info
```


```{r}
if (!file.exists(sst_gd_rds)){
  sst_gd <- griddap(
    sst_info,
    fields    = "sst",
    time      = c("2020-06-16", "2021-06-16"),
    longitude = epu_bb[c("xmin", "xmax")],
    latitude  = epu_bb[c("ymin", "ymax")])
  saveRDS(sst_gd, file = sst_gd_rds)  
}
sst_gd <- readRDS(sst_gd_rds)

sst_gd
```

Now plotting
the griddap object has lots of lists etc in it

Also, getting the coast, an option is in the ecodata package. should check if its only eastern/us though
```{r}
shelf(
  dplyr,
  ggplot2,
  mapdata)

# coastline
coast <- map_data(
  "worldHires", 
  xlim = epu_bb[c("xmin", "xmax")],
  ylim = epu_bb[c("ymin", "ymax")],
  lforce = "e")

sst_df_last <- sst_gd$data %>%
    filter(time == max(time))

# summary(sst_last)

#note that erddap package comes with colors already for standard ocean things
#this is for static plot

ggplot(
  data = sst_df_last, 
  aes(x = lon, y = lat, fill = sst)) +
  geom_polygon(
    data = coast,
    aes(x = long, y = lat, group = group), fill = "grey80") +
  geom_tile() +
  scale_fill_gradientn(
    colors = rerddap::colors$temperature, na.value = NA) +
  theme_bw() + 
  ylab("Latitude") + 
  xlab("Longitude") +
  ggtitle("Latest SST")
```

## extract sst 

we now want to convert data to raster
rasters like to have even widths and heights so we need to round some of the information so that it will play nicely.
Also adjusting time series to have dates and times reasonably
```{r}
shelf(
  purrr,
  raster,
  sp,
  tidyr)
select <- dplyr::select

#clean, round
sst_tbl <- tibble(sst_gd$data) %>%
  mutate(
    # round b/c of uneven intervals
    #   unique(sst_gd$data$lon) %>% sort() %>% diff() %>% table()
    #     0.0099945068359375 0.0100021362304688
    lon  = round(lon, 2),
    lat  = round(lat, 2),
    date = as.Date(time, "%Y-%m-%dT00:00:00Z")) %>%
  select(-time) %>% 
  filter(!is.na(sst)) # 13M to 8.8M rows

#nesting tibbles
#in nest, the not-included column is unique identifier, so here, its date
#also, purrr::map() is going to be 
sst_tbl_mo <- sst_tbl %>% 
  nest(data = c(lat, lon, sst)) %>% 
  mutate(
    raster = purrr::map(data, function(x) {
      #browser() #turning this on will stop execution of code for inspection
      sp::coordinates(x) <- ~ lon + lat
      sp::gridded(x)     <- T
      raster::raster(x)
    }))


#now we have a raster stack
sst_stk <- raster::stack(sst_tbl_mo$raster)
#set names to date
names(sst_stk) <- strftime(sst_tbl_mo$date, "sst_%Y.%m")
#define coordinate reference system for raster object
raster::crs(sst_stk) <- 4326
```


Now we can manipulate our raster stack to do things like, get an average value. etc. then we'll have summarized mean value per month.

However raster::extract outputs a matrix instead of a dataframe. 
row = feature, column = raster layer
To plot this we will need to transform and combine these things.

here we're converting that tough output to tibble, so we're adding columns, pivoting, etc, to get the data to be plottable at all.
We want each series to be a column. so this will be long data with a column 'stat' where we can then filter which stat we are plotting in future commands.

```{r}
shelf(stringr)

#getting raster statistics
epu_sst_avg <- raster::extract(sst_stk, epu_sf, fun = mean, na.rm = T)
epu_sst_sd  <- raster::extract(sst_stk, epu_sf, fun = sd, na.rm = T)

#converting to tibble to be able to use
epu_sst_tbl <- rbind(
  epu_sst_avg %>% 
    as_tibble() %>% 
    cbind(
      EPU  = epu_sf$EPU,
      stat = "mean") %>% 
    pivot_longer(-c(EPU, stat)),
  epu_sst_sd %>% 
    as_tibble() %>% 
    cbind(
      EPU = epu_sf$EPU,
      stat = "sd") %>% 
    pivot_longer(-c(EPU, stat))) %>% 
  mutate(
    EPU  = as.character(EPU),
    date = as.double(str_replace(name, "sst_", ""))) %>% 
  select(-name) %>% 
  pivot_wider(
    names_from  = EPU, 
    values_from = value)
```

now is a plot
```{r}
shelf(dygraphs)

epu_sst_tbl %>% 
  filter(stat == "mean") %>% 
  select(-stat) %>% 
  dygraph()
```


